---
title: "XAI 3: Model-Agnostic methods"
output: html_notebook
---

## EXERCISE:

Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (**cnt**). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.  

## QUESTION:

Analyse the influence of **days since 2011, temperature, humidity** and **wind speed** on the predicted bike counts.


```{r}
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)


#setwd("/Users/cmonserr/OneDrive - UPV/Trabajo_2/Asignaturas/Evaluacion de modelos/Practicas/Practica 3/Bike-Sharing-Dataset")
days <- read.csv("C:/Users/sergi/OneDrive/Escritorio/UPV/3_curso/EDM/Trabajo2/Practica 5 Sin sol/day.csv")
hour <- read.csv("C:/Users/sergi/OneDrive/Escritorio/UPV/3_curso/EDM/Trabajo2/Practica 5 Sin sol/hour.csv")

days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67

rf <- rfsrc(cnt~., data=days_since)

results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)
nr <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

p1 <- ggplot(days_since, aes(x=days_since_2011, y = results$days_since_2011)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Days since 2011") + ylab("Prediction") + ylim(2500,6000)
p2 <- ggplot(days_since, aes(x=temp, y = results$temp)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Temperature") + ylim(2500,6000)
p3 <- ggplot(days_since, aes(x=hum , y = results$hum)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Humidity") + ylim(2500,6000)
p4 <- ggplot(days_since, aes(x=windspeed, y = results$windspeed)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Wind speed") + ylim(2500,6000)

subplot(p1,p2,p3,p4, titleX = TRUE, titleY = TRUE, shareY = TRUE)

```
Comentaremos las gráficas de izquierda a derecha; empezando por Days_since_2011 se obseva que el alquiler de bicis aumenta conforme más tiempo ha pasado desde que la empresa abrió o desde que empezó a guardar datos en la base, la bajada final probablemente se debe a la falta de datos en esa parte. La subida se deve seguramente a que al pasar el tiempo la empresa se ha ido haciendo más conocida, se ha expandido en diferentes lugares, etc.

Ahora vemos la variable Temperature, que según los valores de esta aumentan, más bicis se alquilan: por ejemplo con una temperatura de 10ºC, de media se alquilan unas 3950, pero cuando la temperatura sube hasta los 25ºC se llegan a alquilar 5100 bicis. Esta variable es bastante influyente en el modelo, pues en general, la gente prefiere temperaturas cálidas para dar un paseo en bicicleta. Se ve que a partir de los 28 grados más o menos y cuando aja de los 3 o 4 no hay casi datos por lo que es normal que en esas partes se pierda la tendencia comentada.

Analizando la variable humedad y viento, pasa al contrario que con temperatura, pues probablemente a mayor humedad o mayor viento, la gente tiene menos ganas de coger la bici. Tambien encontramos puntos donde la recta es paralela al eje porque no hay suficientes datos.

## EXERCISE:

Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending of those parameters.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the the data for the Partial Dependency Plot. 

Show the density distribution of both input features with the 2D plot as shown in the class slides. 

TIP: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes. 

## QUESTION:

Interpret the results.


```{r}

sampled <- sample_n(days_since, 40)
temp <- sampled$temp
hum <- sampled$hum
th <- inner_join(data.frame(temp),data.frame(hum), by=character())
th$p <- 0

for(i in 1:nrow(th)){
  r <- days_since
  r[["temp"]] <- th[["temp"]][i]
  r[["hum"]] <- th[["hum"]][i]
  
  sal <- predict(rf, r)$predicted
  th[["p"]][i] <- sum(sal) / nr
}

p5 = ggplot(th, aes(temp, hum, fill = p)) + geom_tile(width = 18, height = 15) + geom_rug() + 
  labs(x="Temperature", y = "Humidity")
p5

```


## EXERCISE:

Apply the previous concepts to predict the **price** of a house from the database **kc_house_data.csv**. In this case, use again a random forest approximation for the prediction based on the features **bedrooms**, **bathrooms**, **sqft_living**, **sqft_lot**, **floors** and **yr_built**. 
Use the partial dependence plot to visualize the relationships the model learned.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot. 

## QUESTION:

Analyse the influence of **bedrooms, bathrooms, sqft_living** and **floors** on the predicted price.


```{r}

d <- read.csv("C:/Users/sergi/OneDrive/Escritorio/UPV/3_curso/EDM/Trabajo2/Practica 5 Sin sol/kc_house_data.csv")

sampled <- sample_n(d, 1000)

sampled <- select(sampled, bedrooms, bathrooms, sqft_living, sqft_lot, floors, yr_built, price)

rf <- rfsrc(price~., data=sampled)

results <- select(sampled, bedrooms, bathrooms, sqft_living, floors, price)
nr <- nrow(sampled)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- sampled
    r[[c]] <- sampled[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

p1 <- ggplot(sampled, aes(x=bedrooms, y=results$bedrooms)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Bedrooms") + ylab("Prediction")
p2 <- ggplot(sampled, aes(x=bathrooms, y=results$bathrooms)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Bathrooms") + ylab("Prediction")
p3 <- ggplot(sampled, aes(x=sqft_living, y=results$sqft_living)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Sqft living") + ylab("Prediction")
p4 <- ggplot(sampled, aes(x=floors, y=results$floors)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Floors") + ylab("Prediction")

subplot(p1,p2,p3,p4, titleX = TRUE)
```

